---
title: "Problema 44"
output: html_document
date: "2024-02-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Problema.** On average 66 % of Finite Mathematics students spend some time in the Mathematics Department's resource room. Half of these students spend more than 90 minutes per week in the resource room. At the end of the semester the students in the class were asked how many minutes per week they spent in the resource room and whether they passed or failed. The passing rates are summarized in the following table:

| Time spent in resource room | Pass % |
|:---------------------------:|:------:|
|            None             |   25   |
|  Between 1 and 90 minutes   |   48   |
|    More than 90 minutes     |   81   |

If a randomly chosen student did not pass the course, what is the probability that he or she did not study in the resource room?

**Solución.** Primero, introduzcamos algo de notación. Tomemos en cuenta los eventos siguientes:

-   $F$: estudiantes del curso de Matemáticas Finitas.

-   $R$: estuvieron en la Sala de Recursos del Departamento de Matemáticas.

-   $N$: no estivieron en la Sala.

-   $L$: estuvieron entre 1 y 90 minutos en la Sala.

-   $M$: estuvieron más de 90 minutos en la Sala.

-   $P$: pasaron el curso.

Entonces, nos solicitan calcular $$P(N\mid\overline P).$$ Por el teorema de Bayes, $$P(N\mid\overline P)=\dfrac{P(\overline P\mid N)P(N)}{P(\overline P)}.$$ De la tabla, podemos obtener $$P(\overline P\mid N)=1 - P(P\mid N)=1-0.25=0.75.$$ También, $$P(N)=1-P(R)=1-0.66=0.34.$$ Ya sólo nos falta calcular $P(\overline P )$. Para ello, calculamos $$\begin{aligned}P(P)&=P(P\cap N) + P(P\cap L) + P(P\cap M)\\
                     &=P(P\mid N) P(N) + P(P\mid L) P(L) + P(P\mid M) P(M)\\
                     &=0.25\cdot0.34+0.48P(L)+0.81P(M).\end{aligned}$$ Entonces, nos falta calcular $P(L)$ y $P(M)$. Obsérvese que $P(M\mid R)=0.5$; por lo que $$P(M)=P(M\cap R) + P(M\cap N)=P(M\mid R)P(R)+0=0.5\cdot0.66=0.33.$$ Como también $P(L\mid R)=0.5$, $P(L)=0.33$ también. Así,$$P(P)=0.25\cdot0.34+0.48\cdot0.33+0.81\cdot0.33.$$ Este valor es

```{r}
0.25 * 0.34 + 0.48 * 0.33 + 0.81 * 0.33
```

Y, entonces, el valor de $P(\overline P)=0.4893$. De esta manera,$$P(N\mid\overline P)=\dfrac{P(\overline P\mid P(N)}{P(\overline P)}=\dfrac{0.75\cdot0.34}{0.4893}.$$ Este valor es

```{r}
0.75 * 0.34 / 0.4893
```
